Prediction of Shared Bicycle Usage at the University of Toronto St. George Campus Using Bayesian Poisson Regression

### 2024-11-28 paper/paper.qmd
readme updated, paper organized
readme updated, paper organized

---
title: "Prediction of Shared Bicycle Usage at the University of Toronto St. George Campus Using Bayesian Poisson Regression"
subtitle: "Usage has risen sharply, with a steep decline in winter, and 79.8% occurs between 8 AM and 8 PM"
author: 
  - Haowei Fan
thanks: "Code and data are available at: https://github.com/HaoweiFan0912/Bikeshare-Forecast.git"
date: today
date-format: long
abstract: "Students and staff at the University of Toronto's St. George campus often face difficulties finding a bike-sharing parking spot when arriving at campus, and finding a bike when leaving, which highlights the need for effective prediction of future bike-sharing demand to support campus commuting infrastructure. This study uses Bayesian Poisson regression to predict the usage of bike-sharing stations at 27 locations on campus for specific years, months, and  four-hour intervals of the day. The results indicate that, on September 26, 2025, from 4:00 to 8:00 AM, usage at 8 stations will increase by 3, 14 stations by 2, 3 stations by 1, 1 station remains unchanged, and 1 station will decrease by 1 compared to the same period in 2024."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction
Bike-sharing services have undoubtedly improved commuting efficiency, especially for short trips within a campus. For example, a study on bike-sharing in Xiamen, China, highlighted how data-driven analyses can inform policymakers about system performance, thereby guiding investments and policy decisions to enhance urban mobility[@xiamen]. However, in the Toronto region, despite a study by the University of Toronto's School of Cities showing that ridership increased from about 665,000 trips in 2015 to over 4.5 million in 2022[@liu2023bike], there has been little research focused on small-scale areas, such as universities, within Toronto. Furthermore, a study from Hangzhou, China, confirmed the significant differences in traveling characteristics between cities and university campuses[@hangzhou]. Therefore, this paper takes the University of Toronto as a case study to analyze bike-sharing usage within the campus.

@@

Structure: TBD

@@

# Model
The model used in this study is as follows:

@@

# Results
This study ultimately predicted the usage volume of 27 shared bicycle stations within the University of Toronto's St. George campus during a 4-hour time window from 4:00 AM to 8:00 AM on September 26, 2025. The prediction was then compared to the usage volume recorded during the same period in 2024. @fig-result shows the results.

@@

# Discussion

>>>>

---
title: "Prediction of Shared Bicycle Usage at the University of Toronto St. George Campus Using Bayesian Poisson Regression"
subtitle: "Usage has risen sharply, with a steep decline in winter, and 79.8% occurs between 8 AM and 8 PM"
author: 
  - Haowei Fan
thanks: "Code and data are available at: https://github.com/HaoweiFan0912/Bikeshare-Forecast.git"
date: today
date-format: long
abstract: "Students and staff at the University of Toronto's St. George campus often face difficulties finding a bike-sharing parking spot when arriving at campus, and finding a bike when leaving, which highlights the need for effective prediction of future bike-sharing demand to support campus commuting infrastructure. This study uses Bayesian Poisson regression to predict the usage of bike-sharing stations at 27 locations on campus for specific years, months, and  four-hour intervals of the day. The results indicate that, on September 26, 2025, from 4:00 to 8:00 AM, usage at 8 stations will increase by 3, 14 stations by 2, 3 stations by 1, 1 station remains unchanged, and 1 station will decrease by 1 compared to the same period in 2024."
format: pdf
toc-depth: 2    
number-sections: true
toc: true           
toc-title: "Table of Contents" 
bibliography: references.bib
---

@@

# Introduction

## Overview
Bike-sharing services have undoubtedly improved commuting efficiency, especially for short trips within a campus. For example, a study on bike-sharing in Xiamen, China, highlighted how data-driven analyses can inform policymakers about system performance, thereby guiding investments and policy decisions to enhance urban mobility[@xiamen]. However, in the Toronto region, despite a study by the University of Toronto's School of Cities showing that ridership increased from about 665,000 trips in 2015 to over 4.5 million in 2022[@liu2023bike], there has been little research focused on small-scale areas, such as universities, within Toronto. Furthermore, a study from Hangzhou, China, confirmed the significant differences in traveling characteristics between cities and university campuses[@hangzhou]. Therefore, this paper takes the University of Toronto as a case study to analyze bike-sharing usage within the campus.

@@

The paper proceeds as follows: @sec-data outlines the data sources, processing techniques, and the variables employed in the study. @sec-model describes the Bayesian Poisson regression model used for prediction, including its formulation and validation. @sec-result discusses the results, emphasizing temporal and station-specific patterns in bike-sharing usage. @sec-discussion concludes with a discussion of the study’s contributions, limitations, and recommendations for future research.

@@

# Model {#sec-model}

## Model descriptions
The model used in this study is as follows:

@@

# Results{#sec-result}
This study ultimately predicted the usage volume of 27 shared bicycle stations within the University of Toronto's St. George campus during a 4-hour time window from 4:00 AM to 8:00 AM on September 26, 2025. The prediction was then compared to the usage volume recorded during the same period in 2024. @fig-result shows the results.

@@

# Discussion{#sec-discussion}
